---
title: "Asian Elephant Occurance Data"
author: "Grace Kumaishi"
date: "1/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse)
library(rgbif)
library(maptools)
library(dismo)
library(rgeos)
library(viridis)
#library(scrubr)
library(raster)
#library(DHARMa) # something is wrong when I try to install/load
library(spocc)
library(sf)
library(rgdal)
library(spData)
library(here)
library(lubridate)
library(kableExtra)

#install wallace
library(wallace)

#load Wallace functions
source(system.file('shiny/funcs', 'functions.R', package = 'wallace'))
```

# This does not need to be run again unless we're changing the input settings to create and thin our occurrence data. All occurrence results are saved as csv's and can be read in with the other Rmd.

## Description:

This markdown file outlines the first steps for creating data for species distribution modeling (SDM) with Wallace.

### Download species occurence data and get coordinates

We begin by downloading the species occurrence points for the Asian Elephant using Wallace's spocc:occ() function. Choose which chunk to run based on the species of interest.

Indian Elephant
```{r}
# Gives global records.  If limit is set high, this function can take a while to run.
results <- spocc::occ(query = "Elephas maximus indicus", # scientific name, here we are using lions
                      from = "gbif", # set to records from https://www.gbif.org/
                      limit = 15000, # max number of records
                      has_coords = TRUE) # gets the lat/long for each observation

# select just GBIF records, format the species name
results[["gbif"]]$data[[formatSpName("Elephas maximus indicus")]]

# select just the necessary information
myspecies_coords <- as.data.frame(results$gbif$data$Elephas_maximus_indicus) %>% # using GBIF data
  dplyr::select(longitude, latitude, occurrenceStatus, coordinateUncertaintyInMeters, institutionCode, references, basisOfRecord, eventDate) %>%  # selecting just the columns that we want
  filter(!basisOfRecord %in% c("FOSSIL_SPECIMEN", "PRESERVED_SPECIMEN", "LIVING_SPECIMEN"),
         occurrenceStatus == "PRESENT",
        eventDate <= "2021-12-31" & eventDate >= "1981-01-01")
```






