---
title: "Asian Elephant Occurrence Data"
author: "Grace Kumaishi"
date: "1/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse)
library(rgbif) # interface to GBIF
library(maptools)
library(dismo)
library(rgeos)
library(viridis)
#library(scrubr)
library(raster)
#library(DHARMa) # something is wrong when I try to install/load
library(spocc) # Interface to species occurrence data sources
library(sf)
library(sp)
library(rgdal)
library(spData)
library(here)
library(lubridate)
library(kableExtra)
library(tmap)
library(here)

#install wallace
library(wallace)

#load Wallace functions
source(system.file('shiny/funcs', 'functions.R', package = 'wallace'))
```

This does not need to be run again unless we're changing the input settings to create and thin our occurrence data. All occurrence results are saved as csv's and can be read in with the other Rmd.

## Description:

This markdown file outlines the first steps for creating data for species distribution modeling (SDM) with Wallace.

### Download species occurence data and get coordinates

We begin by downloading the species occurrence points for the Asian Elephant subspecies using Wallace's spocc:occ() function. Choose which chunk to run based on the species of interest.

Indian Elephant
```{r}
# Gives global records.  If limit is set high, this function can take a while to run.
results <- spocc::occ(query = "Elephas maximus indicus", # scientific name
                      from = "gbif", # set to records from https://www.gbif.org/
                      limit = 15000, # max number of records
                      has_coords = TRUE) # gets the lat/long for each observation

# select just GBIF records, format the species name
results[["gbif"]]$data[[formatSpName("Elephas maximus indicus")]]

# select just the necessary information
indicus_coords <- as.data.frame(results$gbif$data$Elephas_maximus_indicus) %>% # using GBIF data
  dplyr::select(longitude, latitude, occurrenceStatus, coordinateUncertaintyInMeters, institutionCode, references, basisOfRecord, eventDate) %>%  # selecting just the columns that we want
  filter(!basisOfRecord %in% c("FOSSIL_SPECIMEN", "PRESERVED_SPECIMEN", "LIVING_SPECIMEN"),
         occurrenceStatus == "PRESENT",
        eventDate <= "2021-12-31" & eventDate >= "1970-01-01")
```

Sri Lankan Elephant
```{r}
# Gives global records.  If limit is set high, this function can take a while to run.
results <- spocc::occ(query = "Elephas maximus maximus", # scientific name
                      from = "gbif", # set to records from https://www.gbif.org/
                      limit = 15000, # max number of records
                      has_coords = TRUE) # gets the lat/long for each observation

# select just GBIF records, format the species name
results[["gbif"]]$data[[formatSpName("Elephas maximus maximus")]]

# select just the necessary information
maximus_coords <- as.data.frame(results$gbif$data$Elephas_maximus_maximus) %>% # using GBIF data
  dplyr::select(longitude, latitude, occurrenceStatus, coordinateUncertaintyInMeters, institutionCode, references, basisOfRecord, eventDate) %>%  # selecting just the columns that we want
  filter(!basisOfRecord %in% c("FOSSIL_SPECIMEN", "PRESERVED_SPECIMEN", "LIVING_SPECIMEN"),
         occurrenceStatus == "PRESENT",
        eventDate <= "2021-12-31" & eventDate >= "1981-01-01")
```

Sumatran Elephant
```{r}
# Gives global records.  If limit is set high, this function can take a while to run.
results <- spocc::occ(query = "Elephas maximus sumatranus", # scientific name
                      from = "gbif", # set to records from https://www.gbif.org/
                      limit = 15000, # max number of records
                      has_coords = TRUE) # gets the lat/long for each observation

# select just GBIF records, format the species name
results[["gbif"]]$data[[formatSpName("Elephas maximus sumatranus")]]

# select just the necessary information
sumatranus_coords <- as.data.frame(results$gbif$data$Elephas_maximus_sumatranus) %>% # using GBIF data
  dplyr::select(longitude, latitude, occurrenceStatus, coordinateUncertaintyInMeters, institutionCode, references, basisOfRecord, eventDate) %>%  # selecting just the columns that we want
  filter(!basisOfRecord %in% c("FOSSIL_SPECIMEN", "PRESERVED_SPECIMEN", "LIVING_SPECIMEN"),
         occurrenceStatus == "PRESENT",
        eventDate <= "2021-12-31" & eventDate >= "1981-01-01")
```

### Clean Data

```{r}
# remove rows with duplicate coordinates
occs.dups <- duplicated(indicus_coords[c('longitude', 'latitude')])
indicus_coords <- indicus_coords[!occs.dups,]

occs.dups <- duplicated(maximus_coords[c('longitude', 'latitude')])
maximus_coords <- maximus_coords[!occs.dups,]

occs.dups <- duplicated(sumatranus_coords[c('longitude', 'latitude')])
sumatranus_coords <- sumatranus_coords[!occs.dups,]

# make sure latitude and longitude are numeric (sometimes they are characters)
indicus_coords$latitude <- as.numeric(indicus_coords$latitude)
indicus_coords$longitude <- as.numeric(indicus_coords$longitude)

maximus_coords$latitude <- as.numeric(maximus_coords$latitude)
maximus_coords$longitude <- as.numeric(maximus_coords$longitude)

sumatranus_coords$latitude <- as.numeric(sumatranus_coords$latitude)
sumatranus_coords$longitude <- as.numeric(sumatranus_coords$longitude)

# give all records a unique ID
indicus_coords$occID <- row.names(indicus_coords)
maximus_coords$occID <- row.names(maximus_coords)
sumatranus_coords$occID <- row.names(sumatranus_coords)
```

### Crop data by species range

```{r}
# Turn data frames into sf object
indicus_coords_sf <- st_as_sf(x = indicus_coords,
                              coords = c("longitude", "latitude"),
                              crs = 4326)

maximus_coords_sf <- st_as_sf(x = maximus_coords,
                              coords = c("longitude", "latitude"),
                              crs = 4326)

sumatranus_coords_sf <- st_as_sf(x = sumatranus_coords,
                              coords = c("longitude", "latitude"),
                              crs = 4326)

# Check CRS:
# indicus_coords_sf %>% st_crs()
```

```{r}
# Read in shapefile
range <- read_sf(here("redlist_species_data", "data_0.shp"))

# Check crs
#crs(range)

# Crop coordinates by species range file
cropped_indicus_coords <- indicus_coords_sf[range, op = st_intersects]

cropped_maximus_coords <- maximus_coords_sf[range, op = st_intersects]

cropped_sumatranus_coords <- sumatranus_coords_sf[range, op = st_intersects]
```

### Map occurrence data

```{r}
tmap_mode(mode = "view")

asian_elephant_map <-
  tm_shape(cropped_indicus_coords) +
    tm_symbols(col = "red",
               size = 0.2,
               alpha = 0.75) +
  tm_shape(cropped_maximus_coords) +
    tm_symbols(col = "green",
               size = 0.2) +
  tm_shape(cropped_sumatranus_coords) +
    tm_symbols(col = "blue",
                size = 0.2)

asian_elephant_map
```

### Create Asian Continent Shapefile

```{r}
world_sp <- as(world, "Spatial")
world_sf <- st_as_sf(world_sp, "sf")

# filter countries to just those in Asia
asia_countries <- world_sf %>% 
  filter(continent == "Asia") %>% 
  #filter(name_long != )
  dplyr::select(name_long)

# merge countries to create one large polygon
asia <- st_combine(asia_countries)

# convert Asia to spatial polygon
asia_sp <- as_Spatial(asia, cast = TRUE)

indicus_sp <- as_Spatial(cropped_indicus_coords, cast = TRUE)

# check crs
#crs(asia_sp)

intersect <- sp::over(indicus_sp, asia_sp)

# removes na values, so selects for points where they overlap the focal polygon
intersect.rowNums <- as.numeric(which(!(is.na(intersect))))

# filters our occurrence data to those overlapping points
indicus_occs <- cropped_indicus_coords[intersect.rowNums, ] %>% 
  drop_na(longitude, latitude)
indicus_occs$name <- "Elephas maximus indicus" 

# write_csv(indicus_occs, "/Users/gracekumaishi/Desktop/emLab HWC/HWC_RStudio/indicus_points.csv")

write_csv(indicus_occs, "H:/My Drive/Shared with me/HWC_data/Data/Elephants/indicus_points.csv")
```

### Spatial thin

```{r}

```

### Chelsa comparison

```{r}

```




